# Compare means

Calculate mean (SD) and ranges.
Calculate differences in means.
Test differences between cases and controls.

```{r}
rbind(compare("total_area_value", "Total coaptation area, value"),
      compare("orifice_area", "Orifice area"),
      compare("valve_diameter", "Valve diameter"),
      compare("valve_area", "Valve area"),
      compare("a_coap_size_valve", "Total valve coaptation area relative to valve diameter"),
      compare("a_coap_orifice_area", "Total valve coaptation area relative to orifice area"),
      compare("a_coap_valve_area", "Total valve coaptation area relative to valve area"),
      compare("magnitude", "Coaptation line length")) %>%
  mutate(sig = pValue < 0.05) %>% 
  .[names(.) %>% grep("formula", ., invert = TRUE, value = TRUE) %>% c("formula")] %>% 
  select(variable, meanSDCases, meanSDControls, difference, pValue, sig) %>% 
  rename("Variable" = variable,
         "Mean (SD), Cases" = meanSDCases,
         "Mean (SD), Controls" = meanSDControls,
         "Mean difference" = difference,
         "P-value" = pValue,
         "Significance" = sig) %>% 
  kable(digits = 3)
```

## Adjusted comparisons

Adjusted for **orifice area**, `orifice_area`.
After adjusting for orifice area,
the difference in coaptation line lengths of cases and controls is significant.
*Interpretation:*
If we had one case patient and one control patient, both with the same orifice areas,
then their coaptation line lengths would be significantly different.

```{r}
covar <- "orifice_area"
compare("magnitude", "Coaptation line length", covar = covar) %>%
  mutate(sig = pValue < 0.05) %>% 
  .[names(.) %>% grep("formula", ., invert = TRUE, value = TRUE) %>% c("formula")] %>% 
  select(variable, difference, pValue, sig) %>% 
  rename("Variable" = variable,
         "Mean difference" = difference,
         "P-value" = pValue,
         "Significance" = sig) %>% 
  kable(digits = 3)
```
