# Directional analysis of coaptation line

> From: Benjamin Chan   
> Sent: Wednesday, February 08, 2017 8:15 AM  
> To: G. Michael Silberbach <silberbm@ohsu.edu>  
> Cc: Xiao-Yue Han <hanxi@ohsu.edu>  
> Subject: RE: update  
> 
> Hi Michael,
> 
> Yes, I was thinking along the same path as you. The analysis needs to think of
> the "outcome" as the set of the 3 numbers (or 2 numbers, if you want to think
> of the length separately from the angles) that fully describe the centroid
> line. This brings the statistical analysis into what's called a multivariate
> analysis. It's a little more complicated, but possible. Planning on devoting
> some time on Friday to work it out.
> 
> From: G. Michael Silberbach   
> Sent: Tuesday, February 07, 2017 9:29 PM  
> To: Benjamin Chan <chanb@ohsu.edu>  
> Cc: Xiao-Yue Han <hanxi@ohsu.edu>  
> Subject: RE: update  
> 
> Hi Ben,
> 
> Thinking about the centroid line in space, it may be  incorrect to separate it
> into two numbers and compare cases to controls separately.  We are interested
> in comparing the lines not the angles. The two angles (azimuthal and polar )
> separately may not be different from their control separately but a
> combination of the angles could be different.  Also, what about the
> variability in the centroid length AND angle; is it a big cone or a little
> cone? Let's talk about this some more.
> 
> Assuming that we don't need to analyze the centroid angle and length more,
> these data suggest that the sinus mirror hypothesis predicts a normal aortic
> valve in terms of coaptation area and "shape".
> 
> Michael

Correlations between coaptation line measures.

```{r}
df %>%
  select(matches("bsa|^orifice_area$|^(radial|polar|azimuthal)$")) %>%
  cor %>% 
  kable(digits = 2)
```

Scatterplot matrix of coaptation line measures.

```{r scatterplotMatrixCoaptationLine, dpi = 300}
df %>%
  select(matches("^id|^type|^(radial|polar|azimuthal)$")) %>%
  melt(id.vars = c("id", "type")) %>% 
  mutate(type = factor(type)) %>% 
  merge(., ., by = c("id", "type")) %>% 
  filter(variable.x != variable.y) %>% 
  ggplot +
    aes(x = value.x, y = value.y, color = type) +
    geom_point(alpha = 1/2) +
    geom_smooth(method = "lm", se = FALSE) +
    scale_color_brewer("", palette = "Set1") +
    scale_x_continuous("") +
    scale_y_continuous("") +
    facet_grid(variable.y ~ variable.x, scales = "free") +
    theme_bw()
ggsave("../figures/scatterplotMatrixCoaptationLine.svg", dpi = 300)
```

```{r polarplot, dpi = 300}
df %>%
  select(matches("^id|^type|^(polar|azimuthal)$")) %>%
  melt(id.vars = c("id", "type")) %>% 
  mutate(type = factor(type)) %>% 
  merge(., df %>% select(matches("^id|^type|^radial$"))) %>% 
  ggplot +
    aes(x = value, y = radial, color = type, fill = type) +
    geom_point(alpha = 1/2) +
    scale_color_brewer("", palette = "Set1") +
    scale_fill_brewer("", palette = "Set1") +
    scale_x_continuous("", breaks = seq(-180, 180, 30)) +
    scale_y_log10("Length", breaks = c(5, 10, 20)) +
    facet_grid(variable ~ type) +
    coord_polar(start = pi / 2, direction = -1) +
    theme_bw() +
    theme(legend.position = "none")
ggsave("../figures/polarplot.svg", dpi = 300)
```

```{r, echo = TRUE}
x <- matrix(as.numeric(c(df$coapUnitX, df$coapUnitY, df$coapUnitZ)), ncol = 3)
ina <- factor(df$type)
hcf.aov(x, ina)
lr.aov(x, ina)
embed.aov(x, ina)
het.aov(x, ina)
spherconc.test(x, ina)
```

```{r, echo = TRUE}
x <- matrix(as.numeric(c(df$centroidCalcRotatedX - df$centroidBottomRotatedX,
                         df$centroidCalcRotatedY - df$centroidBottomRotatedY,
                         df$centroidBottomZ)),
            ncol = 3)
ina <- factor(df$type)
hcf.aov(x, ina)
lr.aov(x, ina)
embed.aov(x, ina)
het.aov(x, ina)
spherconc.test(x, ina)
```

```{r, echo = TRUE}
x <- matrix(as.numeric(c(df$polar, df$azimuthal) * pi / 180), ncol = 2)
head(x)
ina <- factor(df$type)
hcf.circaov(x, ina, rads = TRUE)
lr.circaov(x, ina, rads = TRUE)
embed.circaov(x, ina, rads = TRUE)
het.circaov(x, ina, rads = TRUE)
conc.test(x, ina, rads = TRUE)
```
